


\documentclass[a4paper]{article}
%\usepackage{a4}
%\usepackage{german}
%\usepackage{ngerman}
%\usepackage{bibgerm}
\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\usepackage{graphics}
\usepackage{psfrag}
\usepackage{hyperref}
%\usepackage{multimedia}
%\usepackage{rotating}
%\usepackage{mathbbol}
%\usepackage{dsfont}
\usepackage{amssymb,amsmath}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{exscale}
\usepackage[round,authoryear]{natbib}
\usepackage{color}
%
% Seitenränder
% \setlength{\unitlength}{1cm} % Laengeneinheit
\setlength{\headsep}{0.5cm} \setlength{\topskip}{0.5cm}
\setlength{\intextsep}{0.1cm} \setlength{\textfloatsep}{0.1cm}
\setlength{\textheight}{24.7cm} \setlength{\topmargin}{-1cm}
\setlength{\textwidth}{17cm} \setlength{\oddsidemargin}{-0.5cm}
\setlength{\evensidemargin}{-0.5cm}
\setlength{\marginparwidth}{50pt} \setlength{\footskip}{1cm}
%
%
% Zeilenabstand
\renewcommand{\baselinestretch}{1.2}\small\normalsize
\input{formel.tex}
\include{fonts}
%
\usepackage{authblk}
\begin{document}

\section{Ciarlet's elastic energy}
The Ciarlet-Geymonat strain energy is close to the St.-Venant-Kirchhoff strain energy, but in contrast to the latter it is polyconvex in $\fF$, 
\begin{align}
w(\fF)=\frac{\lambda}{4}(\III-\ln\III)+\frac{\mu}{2}(\I-\ln\III).
\end{align}
It serves as starting point for our elastic strain gradient extension.
The roman numbers $\I$ and $\III$ denote the first and third principle 
invariants of the Cauchy-Green stretch tensors $\fB$ and $\fC$, 
$\lambda$ and $\mu$ are Lam\'e's constants. 
\section{Strain gradient extension}
To comply with the principle of invariance under rigid body motion, $w$ 
is best formulated in terms of material deformation measures. While we use 
$\fC=\fF^T\fF$ for the first gradient, we need to define a similar 
material stretch measure for the second gradient 
$\fF_3=\fx\otimes\nabla_0\otimes\nabla_0$. A canonical extension is
\begin{align}
\fC_3=\fF^T\fF_3.
\end{align}
A simple invariant is the norm, which we term $g$ to 
indicate the strain gradient contribution,
\begin{align}
g=\sqrt{\fC_3\cdot\cdot\cdot\fC_3}.
\end{align}
We extend $w$ by a summand that contains $g$. We choose
\begin{align}
w_g(\fF)=\frac{\lambda}{4}(\III-\ln\III)+\frac{\mu}{2}(\I-\ln\III)+\frac{\alpha}{2}g^2,
\end{align}
where $\alpha$ is a new material parameter. 
\section{Stress-Strain relation}
We consider the first Piola-Kirchhoff stresses $\fT$ as most practical for 
our purpose. Especially from an implementation point of view, we can 
write down the local balance of momentum by using the material nabla 
operator $\nabla_0$,
\begin{align}
\rho(\fa-\fb)=\fT\cdot\nabla_0-\fT_3\cdot\cdot(\nabla_0\otimes\nabla_0).
\end{align}
Thus we need in the FE implementation the same operators 
$\nabla_0$ and $\nabla_0\otimes\nabla_0$
for calculating the first and second gradient and for 
calculating the residuals, which 
simplifies the implementation considerably. 
We refer to $\fT_3$ thus as the first Piola-Kirchhoff stress tensor of 
third order, which is derived from $w$ similar to the usual first 
Piola-Kirchhoff stresses,
\begin{align}
\fT  &=\frac{\partial w}{\partial \fF},\\
\fT_3&=\frac{\partial w}{\partial \fF_3}.
\end{align}
We summarize $\rho_0$ with the material constants, since it is also a 
constant. The derivatives evaluate to
\begin{align}
\fT  &=\mu \fF + \left(\frac{\lambda}{2}\left(\III-1\right)-\mu\right) 
\fF^{-T}+\alpha F_{mjl} C_{njl}\fe_m\otimes\fe_n,\\
\fT_3 &=\alpha \fF\fC_3=\alpha\fB\fF_3.
\end{align}
The number of indices distinguishes $\fF,\fF_3$ and $\fC,\fC_3$ in the 
index notation. Additional   to the stress-strain (gradient) relation, 
we need the linearisations,
\begin{align}
\frac{\partial\fT}{\partial\fF}   &=\mu \ffI - \left(\frac{\lambda}{2} 
(\III-1)-\mu\right)(F^{-1}_{jk}F^{-T}_{il})_{ijkl}+\lambda\III\fF^{-T}\otimes\fF^{-T}+\alpha(F_{ixy}F_{kxy}\delta_{jl})_{ijkl},\\
\frac{\partial\fT}{\partial\fF_3} &=\alpha(\delta_{ik}C_{jlm}+F_{kj}F_{ilm})_{ijklm},\\
\frac{\partial\fT_3}{\partial\fF} &=\alpha(\delta_{il}C_{mjk}+F_{im}F_{ljk})_{ijklm},\\
\frac{\partial\fT_3}{\partial\fF_3}&=\alpha\fB\ffI^6=\alpha(B_{il}\delta_{jm}\delta_{kn})_{ijklmn}
\end{align}


\end{document}
